# Notes on 'Advanced R'

```{r}
library(pacman)
p_load(magrittr, dplyr)
```


See [web-book](https://adv-r.hadley.nz/vectors-chap.html)

[Pete Wildeford's walkthrough](https://gist.github.com/peterhurford/72dbd44e0a34e29297485a8cf679cf73)

## Integrating stuff from notes


## Vectors

>  for atomic vectors, all elements must have the same type; for lists, elements can have different types. 


> Every vector can also have attributes, which you can think of as a named list of arbitrary metadata. Two attributes are particularly important. The dimension attribute turns vectors into matrices and arrays and the class attribute powers the S3 object system. 



## Aspects of R, pros and cons

Some of the best features are:

> Powerful metaprogramming facilities. R is not just a programming language, it is also an environment for interactive data analysis. Its metaprogramming capabilities allow you to write magically succinct and concise functions and provides an excellent environment for designing domain-specific languages


> Much of the R code you-ll see in the wild is written in haste to solve a pressing problem.

> has aspects of both functional and object-oriented (OO) programming languages

## Data structures

> R-s base data structures can be organised by their dimensionality (1d, 2d, or nd) and whether they-re homogeneous (all contents must be of the same type) or heterogeneous (the contents can be of different types). This gives rise to the five data types most often used in data analysis  (p33)

Homogeneous, Heterogeneous
1d Atomic vector, List
2d Matrix, Data frame
nd Array


> Note that R has no 0-dimensional, or scalar types. Individual numbers or strings, which you might think would be scalars, are actually vectors of length one.

> factors, an important data structure created by setting attributes of an atomic vector.


> Atomic vectors are usually created with c(), short for combine:

```{r}

(
  dbl_var <- c(1, 2.5, 4.5) 
) %T>% 
  str() 

```

>  With the L suffix, you get an integer rather than a double

```{r}

int_var <- c(1L, 6L, 10L) %T>% str()

```

> Use TRUE and FALSE (or T and F) to create logical vectors

Quotes for characters

```{r}

log_var <- c(TRUE, FALSE, T, F) %T>% 
  str()

chr_var <- c("these are", "some strings") %T>% 
  str()


```


> Atomic vectors are always flat, even if you nest c()-s:


`c(1, c(2, c(3, 4)))`

`r c(1, c(2, c(3, 4)))`

 [1] 1 2 3 4
 
> is  the same as `c(1, 2, 3, 4)`

`r c(1, 2, 3, 4)`

>  when you attempt to combine different types they will be coerced to the most flexible type. Types from least to most flexible are: logical, integer, double, and character

> When a logical vector is coerced to an integer or double, TRUE becomes 1 and FALSE becomes 0. This is very useful in conjunction with sum() and mean()

```{r}

x <- c(FALSE) %T>%
  str()

```

(p37)
You construct lists by using list() instead of
c():
x <- list(1:3, "a", c(TRUE, FALSE, TRUE), c(2.3, 5.9))
str(x)
#> List of 4
#> $ : int [1:3] 1 2 3
#> $ : chr "a"
#> $ : logi [1:3] TRUE FALSE TRUE
#> $ : num [1:2] 2.3 5.9
Lists are sometimes called recursive vectors

(p38)
You can turn a list into an atomic vector
with unlist(). If

(p38)
Lists are used to build up many of the more complicated data structures
in R. For example, both data frames (described in Section 2.4) and linear
models objects (as produced by lm()) are lists

(p39)
-??
Why is the default missing value, NA, a logical vector? What-s
special about logical vectors? (Hint: think about c(FALSE,
NA_character_).)

(p39)
All objects can have arbitrary additional attributes, used to store metadata about the object. Attributes can be thought of as a named list (with
unique names). Attributes can be accessed individually with attr() or
all at once (as a list) with attributes().

(p40)
By default, most attributes are lost when modifying a vector:
attributes(y[1])
#> NULL
attributes(sum(y))
#> NULL
The only attributes not lost are the three most important:
- Names, a character vector giving each element a name, described in
Section 2.2.0.1.
- Dimensions, used to turn vectors into matrices and arrays, described
in Section 2.3.
- Class, used to implement the S3 object system, described

(p40)
Each of these attributes has a specific accessor function to get and set
values. When working with these attributes, use names(x), class(x), and
dim(x), not attr(x, "names"), attr(x, "class"), and attr(x, "dim

(p40)
You can name a vector in three ways

(p40)
name. If some names are
missing, names() will return an empty string for those elements

(p41)
A factor is a vector
that can contain only predefined values, and is used to store categorical
data. Factors are built on top of integer vectors using two attributes:
the class(), -factor”, which makes them behave differently from regular
integer vectors, and the levels(), which defines the set of allowed values

(p39)
Attributes can be accessed individually with attr() or
all at once (as a list) with attributes()

(p41)
-levels set based on initial assignment
x <- factor(c("a", "b", "b", "a"))

(p41)
# You can't use values that are not in the levels
x[2] <- "c"
#> Warning: invalid factor level, NA generated

(p42)
-making a character vector into a factor
sex_char <- c("m", "m", "m")
sex_factor <- factor(sex_char, levels = c("m", "f"))

(p42)
# Or change how we read it in:
z <- read.csv(text = "value\n12\n1\n.\n9", na.strings=".")
typeof(z\$value)


(p43)
Unfortunately, most data loading functions in R automatically convert
character vectors to factors. This is suboptimal, because there-s no
way for those functions to know the set of all possible levels or their
optimal order. Instead, use the argument stringsAsFactors = FALSE
to suppress this behaviour, and then manually convert character vectors to factors using your knowledge of the data

(p43)
-s usually best
to explicitly convert factors to character vectors if you need string-like
behaviour.

(p42)
makes it obvious when some groups contain
no observations:
sex_char <- c("m", "m", "m"

(p39)
The structure() function returns a new object with modified attributes

(p40)
structure(1:10, my_attribute = "This is a vector")

(p43)
Some
string methods (like gsub() and grepl()) will coerce factors to strings,
while others (like nchar()) will throw an error, and still others (like c())
will use the underlying integer values.

(p44)
Matrices and arrays are created with matrix() and array(), or by using
the assignment form of dim():
# Two scalar arguments to specify rows and columns
a <- matrix(1:6, ncol = 3, nrow = 2)
# One vector argument to describe all dimensions
b <- array(1:12, c(2, 3, 2))
# You can also modify an object in place by setting dim()
c <- 1:6
dim(c) <- c(3, 2)

(p45)
length(a)
#> [1] 6
nrow(a)
#> [1] 2
ncol(a)
#> [1] 3
rownames(a) <- c("A", "B")
colnames(a) <- c("a", "b", "c")

(p45)
c() generalises to cbind() and rbind() for matrices

(p45)
You can transpose a matrix with
t()

(p46)
str(matrix(1:3, ncol = 1)) # column vector
#> int [1:3, 1] 1 2 3

(p46)
the dimension attribute can also be set on lists to make list-matrices or listarrays:

(p47)
2.4 Data frames

(p47)
Under the hood, a data frame is a list of
equal-length vectors

(p48)
-??
Because a data.frame is an S3 class, its type reflects the underlying
vector used to build it: the list. To check

(p47)
Beware data.frame()-s default behaviour which turns strings into factors.
Use stringAsFactors = FALSE to suppress this behaviour:
df <

(p48)
-class vs type
class

(p48)
You can coerce an object to a data frame with as.data.frame():

(p48)
You can combine data frames using cbind() and rbind(

(p49)
Use plyr::rbind.fill() to combine
data frames that don-t have the same columns

(p49)
Since a data frame is a list of vectors, it is possible for a data frame to
have a column that is a list

(p50)
-what's I()?
A workaround is to use I(), which causes data.frame() to treat the list
as one unit:
dfl <- data.frame(x = 1:3, y = I(list(1:2, 1:3, 1:4)))

(p50)
Similarly, it-s also possible to have a column of a data frame that’s a
matrix or array, as long as the number of rows matches the data frame

(p50)
-but why?
Similarly, it-s also possible to have a column of a data frame that’s a
matrix or array

(p54)
There are five things that you can use to subset a vector:
- Positive integers return elements at the specified positions:
x[c(3, 1)]
#> [1] 3.3 2.1

(p54)
Negative integers omit elements at the specified positions:
x[-c(3, 1)]
#> [1] 4.2 5.4

(p54)
Logical vectors select elements where the corresponding logical value
is TRUE. This is probably the most useful type of subsetting because
you write the expression that creates the logical vector:
x[c(TRUE, TRUE, FALSE, FALSE)]

(p54)
Logical vectors select elements where the corresponding logical value
is TRUE. This is probably the most useful type of subsetting because
you write the expression that creates the logical vector:
x[c(TRUE, TRUE, FALSE, FALSE)]
#> [1] 2.1 4.2
x[x > 3]
#> [1] 4.2 3.3 5.4
If the logical vector is shorter than the vector being subsetted, it will
be recycled to be the same length.
x[c(TRUE, FALSE)]
#> [1] 2.1 3.3
# Equivalent to
x[c(TRUE, FALSE, TRUE, FALSE)]
#> [1] 2.1 3.3

(p55)
If the vector is named, you can also use:
- Character vectors to return elements with matching names.
(y <- setNames(x, letters[1:4]))
#> a b c d
#> 2.1 4.2 3.3 5.4
y[c("d", "c", "a")]
#> d c a
#> 5.4 3.3 2.1
#

(p55)
-letters [] is cool
letters

(p56)
Lists

(p56)
Using [ will always return a list

(p56)
a[1:2, ]
#> A B C
#> [1,] 1 4 7
#> [2,] 2 5 8

(p56)
-??
Arrays in R are stored in column-major order:
(vals <- outer(1:5, 1:5, FUN = "paste", sep = ","))
#> [,1] [,2] [,3] [,4] [,5]

(p57)
 if you
subset with a single vector, they behave like lists; if you subset with two
vectors, they behave like matrices

(p58)
# Like a list:
df[c("x", "z")]
#> x z
#> 1 1 a
#> 2 2 b
#> 3 3 c
# Like a matrix
df[, c("x", "z")]

(p57)
df[df$x == 2, ]
#> x y z
#> 2 2 2 b

(p58)
-simplifies means reducing it to a vector
 There's an important difference if you select a single
# column: matrix subsetting simplifies by default, list
# subsetting does not.
str(df["x"])
#> 'data.frame': 3 obs. of 1 variable:
#> $ x: int 1 2 3
str(df[, "x"])
#> int [1:3] 1 2 3

(p58)
-what's s3 vs s4 objects?
5 S3 objects

(p59)
[ is similar to [,
except it can only return a single value and it allows you to pull pieces
out of a list. $ is a useful shorthand for [[ combined with character
subsetting.
You need [

(p59)
[[ is similar to [,
except it can only return a single value and it allows you to pull pieces
out of a list.

(p59)
-If list x is a train carrying objects, then x[[5]] is the object in car
5; x[4:6] is a train of cars 4-6.-

(p60)
a <- list(a = 1, b = 2)
a[[1]]
#> [1] 1
a[["a"]]
#> [1] 1

(p60)
Because data frames are lists of columns, you can use [[ to extract a
column from data frames: mtcars[[1]], mtcars[["cyl"]]

(p60)
Simplifying subsets returns the simplest possible
data structure that can represent the output, and is useful interactively
because it usually gives you what you want. Preserving subsetting keeps
the structure of the output the same as the input, and is generally better
for programming because the result will always be the same type

(p60)
Unfortunately, how you switch between simplifying and preserving differs for different data types, as summarised in the table below

(p61)
Preserving is the same for all data types: you get the same type of
output as input. Simplifying behaviour varies slightly between different
data types, as described below

(p61)
Data frame x[, 1] or x[[1]] x[, 1, drop = F] or x[1]

(p61)
Factor: drops any unused levels.

(p61)
Matrix or array: if any of the dimensions has length 1, drops that
dimension.

(p62)
Data frame: if output is a single column, returns a vector instead of
a data frame

(p62)
$ is a shorthand operator, where x$y is equivalent to x[["y", exact
= FALSE]]. It-s often used to access variables in a data frame, as in
mtcars$cyl or diamonds$carat.

(p62)
-seems useful
One common mistake with $ is to try and use it when you have the name
of a column stored in a variable:
var <- "cyl"
# Doesn't work - mtcars$var translated to mtcars[["var"]]
mtcars$var
#> NULL
# Instead use [[
mtcars[[var]]
#> [1] 6 6 4 6 8 6 8 4 4 6 6 8 8 8 8 8 8 4 4 4 4 8 8 8 8 4 4 4 8
#> [30] 6 8 4

(p62)
$ does partial matching

(p64)
x[c(1, 2)] <- 2:3

(p64)
[-1] <- 4:1

(p65)
Subsetting with nothing can be useful in conjunction with assignment
because it will preserve the original object class and structure. Compare
the following two expressions. In the first, mtcars will remain as a data
frame. In the second, mtcars will become a list.
mtcars[] <- lapply(mtcars, as.integer)
mtcars <- lapply(mtcars, as.integer)

(p33)
str() is short for structure and it gives a
compact, human readable description of any R data structure

(p34)
The basic data structure in R is the vector

(p34)
Type, typeof(), what it is

(p35)
Length, length(), how many elements it contains.
- Attributes, attributes(), additional arbitrary metadata

(p35)
all elements of an atomic
vector must be the same type, whereas the elements of a list can have
different types.

(p36)
Given a vector, you can determine its type with typeof(), or check if
it-s a specific type with an “is” function: is.character(), is.double(),
is.integer(), is.logical(), or, more generally, is.atomic()

(p37)
-!! list vs atomic vector
including lists. You construct lists by using list() instead of
c():

(p37)
Lists are sometimes called recursive vectors, because a list can contain other lists

(p38)
-linear model objects are lists
and linear
models objects (as produced by lm()) are lists

(p42)
na.strings argument to read.csv(

(p42)
-to test things
Reading in "text" instead of from a file here:
z <- read.csv(text = "value\n12\n1\n.\n9")

(p42)
-?
so why is it calling it a integer here if it's turning it into a factor?
typeof(z$value)
#> [1] "integer"

(p42)
-?l typeof vs class functions?
class(z$value)
#> [1] "factor"

(p47)
You create a data frame using data.frame(), which takes named vectors
as input:
df <- data.frame(x = 1:3, y = c("a", "b", "c"))
str(df)
#> 'data.frame': 3 obs. of 2 variables:
#> $ x: int 1 2 3
#> $ y: Factor w/ 3 levels "a","b","c": 1 2 3

(p59)
$ is a useful shorthand for [[ combined with character
subsetting.
