# Getting, cleaning and using data {#data-sci}

This will build on my content [here](https://daaronr.github.io/writing_econ_research/economic-theory-modeling-and-empirical-work.html#getting-and-using-data), and integrate with it.

Some key resources are in a continually updated airtable [HERE](https://airtable.com/shrqNt0YAa3eLiK5S)

See especially:

[R for data science](r4ds.had.co.nz)

[Advanced R](https://adv-r.hadley.nz/)

[bookdown: Authoring Books and Technical Documents with R Markdown:](https://bookdown.org/yihui/bookdown/)

[OSF: 'PhD Toolkit on Transparent, Open, Reproducible Research'	https://osf.io/g8yjz/](OSF: 'PhD Toolkit on Transparent, Open, Reproducible Research'	https://osf.io/g8yjz/)


[Happy Git and GitHub for the useR](https://happygitwithr.com/)


"Data science for business" see [WIP notes here](#n_ds4bs)


\


"Code and Data for the Social Sciences" (Gentzkow/Shapiro)


## Data: What/why/where/how

## Organizing a project

## Dynamic documents (esp Rmd/bookdown)

Some guidelines from a particular project:

[Appendix: Tech for creating, editing and collaborating on this ‘Bookdown’ web book/project (and starting your own)](https://daaronr.github.io/ea_giving_barriers/bookdown-appendix.html)

### Managing references/citations


```{block2,  type='note'}
A letter to my co-authors...

Hi all.

Hope you are doing well. I've just invited you to a shared Zotero group managing my bibliography/references. I think this should be useful. (I prefer Zotero to Mendeley because it's open source and... I forgot the other reason.)
On my computer it synchronizes with a .bib (bibtex) file in a dropbox folder ...

For latex files we just refer to this as normal.
In the Rmd files/bookdown (producing output like [EA barriers](#https://daaronr.github.io/ea_giving_barriers/outline.html) or Metrics notes (present book) this is referenced in the YAML header to the index.Rmd file as

> bibliography: [reinstein_bibtex.bib]

Then, to keep this file, I have a "download block" included in that same file (the first line with 'dropbox' is the key one).

```

\

The download code follows (remove the 'eval=FALSE' to get it to actually run)...

```{r download-formatting-to-support-folder-example, eval=FALSE}

tryCatch( #trycatch lets us 'try' to execute and if there is an error, it does the thing *after* the braces, rather than crashing
  {
   download.file(url = "https://www.dropbox.com/s/3i8bjrgo8u08v5w/reinstein_bibtex.bib?raw=1", des tfile = "reinstein_bibtex.bib") #download the bibtex database

        download.file(url = "https://raw.githubusercontent.com/daaronr/dr-rstuff/master/bookdown_template/support/tufte_plus.css", destfile = here("support", "tufte_plus.css")) #this downloads the style file

  }, error = function(e) {
    print("you are not online, so we can't download")
  }
)
```


A fairly comprehensive discussion of tools for citation in R-markdown:

[A Roundup of R Tools for Handling BibTeX](https://ropensci.org/technotes/2020/05/07/rmd-citations/)


### An example of dynamic code


Shapiro Wilk test for normality; professor salaries at some US university from the built in Cars data...

<div class="marginnote">
By the way, if anyone wants me to offer me a job at that university, it looks like a great deal!
</div>

```{r Diagnostic-tests, include=TRUE}


prof_sal_shapiro_test <- shapiro.test(carData::Salaries$salary)

  # ShapiroTest <- map_df(list(SXDonShapiroTest, EXDonShapiroTest), tidy)
  # (ShapiroTest <- kable(ShapiroTest) %>% kable_styling())
```

The results from the Shapiro Wilk normality test ...

The p-values are  `r prof_sal_shapiro_test[2]` suggesting this data `r ifelse(prof_sal_shapiro_test[2]<0.10, "is not", "is")` normal

## Project management tools, esp. Git/Github

(More to be added/linked here)

See ['Git and GitHub' here... watch this space](https://daaronr.github.io/ea_giving_barriers/bookdown-appendix.html#git-and-github)

\

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">For students and research assistants, I&#39;ve been sending first time git users/developers to this:<a href="https://t.co/P6KQpXHCWI">https://t.co/P6KQpXHCWI</a><br>+ <a href="https://t.co/q4R4Ei5Biw">https://t.co/q4R4Ei5Biw</a></p>&mdash; Nathan Lane (\@straightedge) <a href="https://twitter.com/straightedge/status/1172694350205087744?ref_src=twsrc%5Etfw">September 14, 2019</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

## Good coding practices

### New tools and approaches to data (esp 'tidyverse')

From [Kurtz](https://bookdown.org/content/3890/small-worlds-and-large-worlds.html):

> If you are new to tidyverse-style syntax, possibly the oddest component is the pipe (i.e., `%>%`). I'm not going to explain the `%>%` in this project, but you might learn more about in [this brief clip](https://www.youtube.com/watch?v=9yjhxvu-pDg), starting around [minute 21:25 in this talk by Wickham](https://www.youtube.com/watch?v=K-ss_ag2k9E&t=1285s), or in [section 5.6.1 from Grolemund and Wickham's *R for Data Science*](https://r4ds.had.co.nz/transform.html#combining-multiple-operations-with-the-pipe). Really, all of Chapter 5 of *R4DS* is just great for new R and new tidyverse users. And *R4DS* Chapter 3 is a nice introduction to plotting with ggplot2.

### Style and consistency

#### lower_snake_case {-}

Use `lower_snake_case` to name *all* objects (that's my preference anyways)
unless there's a strong reason to do otherwise.

\

This includes:

`file_names.txt`
`folder_names`
`function_names` (with few exceptions)
`names_of_data_objects_like_vectors`
`names_of_data_output_objects_like_correlation_coefficients`
`ex_df1` In R you probably should keep data frame names short to avoid excessive typing

\

*And by all that is holy, never put spaces or slashes in file or object names!* This can make it very hard to process across systems... there are various ways of referring to spaces and other white space.

#### Indenting and spacing


### Using functions, variable lists, etc., for clean, concise, readable code

### Mapping over lists to produce results



<!-- ft <- list(ft_treat_no_ask, ft_treat_no_ask_19, ft_treat_no_ask_19_short, ft_treat_no_ask_19_long)

ft_names <- c("All", "2019", "2019-short", "2019-long")

ft <- map2(ft, ft_names, function(x, y) {
  broom::tidy(x) %>% add_column(Experiment = y)
}) %>%
  bind_rows() %>%
  kable(, caption = "S2 Donation incidence by S1 Ask/no-ask; Fisher tests", digits = 2) %>%
  kable_styling()
-->

### Building results based on 'lists of filters' of the data set

In writing a paper you very often want to produce 'statistics A-F for subsets S1-S5'.

Can't we just automate this?

\


We can store a filter as a character vector and then apply it
```{r}

selection_statement <- "Species == 'setosa' & Petal.Width>0.3"

iris %>%
   as.tibble() %>%
    filter(rlang::eval_tidy(rlang::parse_expr(selection_statement)))
```

Making this a function for later use:

```{r}

selection_statement <- "Species == 'setosa' & Petal.Width>0.3"


filter_parse =  function(df, x) {
  {{df}} %>%
   filter(rlang::eval_tidy(rlang::parse_expr({{x}})))
}

iris %>%
  as.tibble() %>%
 filter_parse(selection_statement)

```



We can do the same for a list of character vectors of filter statements, and apply each filter from the list to the dataframe, and then the output function:

```{r}

sel_st <- c("Species == 'setosa' & Petal.Width>0.3", "Species == 'virginica' & Petal.Width>2.4")

map(iris, selection_statement)

sel_st %>% map(~ filter_parse(iris, .x))

```

It works nicely if you have a list of filters aligned with a list of names or other objects 'specific to each filter'


(Code below: adapt to public data and explain)

```
 bal_sx <-
    map2(subsets_sx_dur, subsets_sx_dur_name, function(x, y) {
      filter_parse(sa, x) %>%
    dplyr::filter(stage == "2") %>%
     tabyl(treat_1, treat_2, show_missing_levels = FALSE)  %>%
    adornme_not(cap = paste(y, "; 1st and 2nd stage treatments"))
    }
  )
```


### Coding style and indenting in Stata (one approach)

I indent every line except

- clear, import, save, merge ('file operations')
  - except where these occur as part of a loop, in which case I put in an 'important comment' noting these operations

- lines that call other do files

- important comments/flags/to-do's


I only put 'small todo' elements having to do with code in a code file itself (and even then there may be better places). If we are going to put todos I suggest we include `#todo` to search for these later (and R has a utility to collect these in a nice way... maybe Stata does too.


\


Whenever there are more than 20 lines of something prosaic that cannot/has not been put into a loop or function, I suggest we put it in a separate 'do' file and call that do file (with no indent).  That's what I do here, giving a brief description and a 'back link'.  Sometimes I put all those do files into an separate folder.


## Additional tips (integrate)

[When you have to upgrade R on Mac, how to preserve package installations - twitter thread](https://twitter.com/gdequeiroz/status/1228722821817225216)


- [This worked well for me](https://github.com/ivelasq/r-data-recipes/blob/master/README.md#reinstall-packages-after-a-major-r-update). Thanks
\@ivelasq3 !
\

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">Are you teaching/learning <a href="https://twitter.com/hashtag/r?src=hash&amp;ref_src=twsrc%5Etfw">#r</a> <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> &amp; want to teach/learn the latest <a href="https://twitter.com/hashtag/tidyverse?src=hash&amp;ref_src=twsrc%5Etfw">#tidyverse</a> <a href="https://twitter.com/hashtag/tidyr?src=hash&amp;ref_src=twsrc%5Etfw">#tidyr</a> tools? e.g. bind_rows, pivot_wider/pivot_longer, the join family (full_join, inner_join...) Check out my slides on &quot;Advanced data manipulation&quot; here 😺<a href="https://t.co/dr9VNx7MFf">https://t.co/dr9VNx7MFf</a></p>&mdash; Amy Willis (\@AmyDWillis) <a href="https://twitter.com/AmyDWillis/status/1195498569072959490?ref_src=twsrc%5Etfw">November 16, 2019</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

\

**Tools for structuring your workflow for reproducable code with Rmd and Git: The workflowr package**

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">&quot;This paper does a thorough job setting out the rationale, design, and implementation of the workflowr package&quot; says <a href="https://twitter.com/PeteHaitch?ref_src=twsrc%5Etfw">\@PeteHaitch</a> <a href="https://twitter.com/WEHI_research?ref_src=twsrc%5Etfw">\\@WEHI_research</a> in his review of this <a href="https://twitter.com/hashtag/softwaretool?src=hash&amp;ref_src=twsrc%5Etfw">#softwaretool</a> article introducing workflowr by <a href="https://twitter.com/jdblischak?ref_src=twsrc%5Etfw">\@jdblischak</a> and co-authors <a href="https://t.co/ZXmQkDhFuD">https://t.co/ZXmQkDhFuD</a> <a href="https://twitter.com/hashtag/OpenScience?src=hash&amp;ref_src=twsrc%5Etfw">#OpenScience</a> <a href="https://t.co/e77SVo8PhO">pic.twitter.com/e77SVo8PhO</a></p>&mdash; F1000Research (\@F1000Research) <a href="https://twitter.com/F1000Research/status/1196056651691962368?ref_src=twsrc%5Etfw">November 17, 2019</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

\

Switching from Latex to markdown/R-markdown? [These tips from Colin Bousige look pretty good](https://colinbousige.github.io/post/rmarkdown/), although I prefer the bookdown/gitbook format


Massive data cleaning using the Recipe package


Codebook package
